Hierarchically Supervised Latent Dirichlet Allocation

### 2011-05-25 paper/main.tex
added a few words
Updated Experiments

\subsection{Data}

Our data set was gathered from the clinical data warehouse of NewYork
- Presbyterian Hospital. The data consisted of free-text discharge
summaries and their respective ICD-9 codes. A discharge summary is
a clinical report prepared by a physician or other health professional
at the conclusion of a hospital stay or series of treatments. The
note outlines the patient\textquoteright{}s chief complaint, diagnostic
findings, therapy administered, patient\textquoteright{}s response
to the chosen therapy, the treatment plan and the recommendations
upon discharge. The ICD-9 codes used to structure the discharge summary
data are part of a controlled terminology which is the international
standard diagnostic classification for epidemiological, health management,
and clinical purposes (http://www.who.int/classifications/icd/en/).
The codes are classified in a rooted-tree structure, with each edge
representing an is-a relationship between parent and child, such that
the parent diagnosis subsumes the child diagnosis. In the hospital,
ICD-9 codes are generated manually by trained medical coders, who
review all the information in the discharge summary. For the purposes
of sLDA, ICD-9 codes will be used as labels for discharge summaries.

We worked within the guidelines of the Health Insurance Portability
and Accountability Act (HIPAA), which protects patient privacy and
the security of potentially identifying medical material, known as
personal health information (PHI). HIPAA covers any information within
a medical record that was created, used, or disclosed during the course
of providing a health care service and that can be used to identify
an individual. Before beginning data processing, we generated a PHI-free
dataset (see Data Pre-processing below).

\subsection{Pre-Processing}

Patient discharge summaries and their associated ICD-9 diagnoses are
stored in two different places in the NewYork - Presbyterian data
warehouse, and so had to be linked together before being fed into
the sLDA algorithm. Each discharge note and set of diagnoses were
assigned a patient unique identifier (PUID) and a visit unique identifier
(VUID), allowing the two types of data to be linked.

Natural Language Processsing (NLP) techniques were used to process
the free-text discharge summaries. First, the Natural Language Toolkit
(http://www.nltk.org/) was used to tokenize the text. Next, feature
selection was performed using a term frequency - inverse document
frequency (TF-IDF) algorithm on the entire document set and sorting
the words by their TF-IDF values. The top 10,000 words were manually
evaluated to eliminate all potentially identifying information. Finally,
each discharge summary was converted to a bag-of-words, listing the
frequencies of the remaining, free of protected health information,
top 10,000 words.

Preparation of the diagnostic codes involved inference over the ICD-9
hierarchy. The is-a relationships of the hierarchy allowed us to make
two important assumptions. First, if a diagnosis was observed to be
present, all of its ancestors could be assumed to be present as well
(e.g., if a patient had malignant hypertension, it could be assumed
that they also had essential hypertension. Second, if a diagnosis
was observed to be absent, it could be assumed that all of its descendants
were also absent (e.g. if a patient did not have essential hypertension,
it could be assumed that they did not have malignant hypertension).
Unfortunately, ICD-9 code observations never include observations
of disease absence. ICD-9 codes are only documented when the condition
is observed to be present. Additionally, ICD-9 codes are known to
have relatively low sensitivity; conditions that are present are often
not documented in a set of ICD-9 codes (Surjan, 1999). Given these
facts, we made the following assumptions regarding each visit: recorded
diagnoses and their ancestors were labeled as true; diagnoses that
were observed at some time for a patient but not at the current visit
were labeled as unobserved; and diagnoses that had never been listed
for a patient were labeled as false for all of that patient\textquoteright{}s
visits. This last assumption captures the belief that parts of the
ICD-9 hierarchy that are never observed for a particular patient are
almost certain to be false. Additionally, for computational purposes,
we decided not to include an ICD-9 code at all if neither it nor one
of its descendants had been assigned to a patient in any of the records
in our dataset.

\subsection{ICD-9 Code Hierarchy}

Here, we augment the sLDA model such that the supervised signal is
distribution over the ICD-9 code tree, which is an is-a hierarchy
\citep{ICD9NCBO}. An is-a hierarchy is represented by the tree data
structure where each node has only a single parent and nodes cannot
be parents of ancestors (ie. there are no loops). In this particular
case, the ICD-9 code hierarchy is also partially a prefix trie where
the labels for certain nodes are prefixes for child nodes. Given that
this rule does not apply to all nodes in the hierarchy, we did not
use this feature to determine the structure of the hierarchy. Instead
we acquired a dataset that explicitly defined the relationships between
the nodes of the hierarchy \citep{ICD9NCBO}. In documentation of
ICD-9 codes for billing purposes, only a subset of the nodes can be
used, however the nodes higher in the hierarchy contain semantic information
about the categories of codes that are their descendants. For this
reason, we included these nodes in our model.

>>>>

In the section we describe the application of HSLDA for prediction
in two hierarchically structured domains. Firstly, we describe using
discharge summaries to predict diagnoses, encoded as ICD-9 codes.
Discharge summaries are documents that are authored by clinicians
to summarize the course of someone who has been hospitalized. ICD-9
codes are used mainly for billing purposes to indicate the conditions
for which a patient was treated. Secondly, we describe using Amazon.com
product descriptions to predict product categories.

\subsection{Diagnosis Prediction}

Our data set was gathered from the clinical data warehouse of NewYork-Presbyterian
Hospital. The data consisted of free-text discharge summaries and
their respective ICD-9 codes. A discharge summary is a clinical report
prepared by a physician or other health professional at the conclusion
of a hospital stay or series of treatments. The note outlines the
patient's chief complaint, diagnostic findings, therapy administered,
patient's response to the chosen therapy, the treatment plan and
the recommendations upon discharge. The ICD-9 codes used to structure
the discharge summary data are part of a controlled terminology which
is the international standard diagnostic classification for epidemiological,
health management, and clinical purposes. The ICD-9 codes are organized
in a rooted-tree structure, with each edge representing an is-a relationship
between parent and child, such that the parent diagnosis subsumes
the child diagnosis. For example, the code representing {}``Pneumonia
due to adenovirus'' is a child of the code representing {}``Viral
pneumonia'' where the former is a type of the latter. In the hospital,
ICD-9 codes are generated manually by trained medical coders, who
review all the information in the discharge summary.

The text of the discharge summaries were pre-processed such that each
document would be represented as counts over a 10,000 word vocabulary.
The Natural Language Toolkit was used to tokenize the text. A vocabulary
was identified by first sorting terms based on a global term frequency-inverse
document frequency measure. The top 10,000 words which were not identifying
in some way (a name, place, or identifying number) were selected for
inclusion in the vocabulary.

For each hospitalization there are usually several ICD-9 codes assigned
for billing purposes. These codes are known to be quite specific but
not very sensitive {[}citation{]}. Regardless of that fact, this is
one of the only sources for information on patient diagnoses aside
from the free text. Aside from prediction, one of the goals is to
compare the sensitivity of predictions from the HSLDA model in comparison
to the codes in a case where a test closer to ground truth is available.
For this we will compare whether predictions for the ICD-9 code associated
with anemia are better predicted by HSLDA or by the ICD-9 codes. Anemia
was chosen because hemoglobin values are readily available and the
definition of anemia according the World Health Organization is approximately
12.5, with a threshold of 12 for women and 13 for men {[}citation{]}.

We worked within the guidelines of the Health Insurance Portability
and Accountability Act (HIPAA), which protects patient privacy and
the security of potentially identifying medical material, known as
personal health information (PHI). HIPAA covers any information within
a medical record that was created, used, or disclosed during the course
of providing a health care service and that can be used to identify
an individual. This study was approved by the Institutional Review
Board.

\subsection{Product Category Prediction}

Data for these experiments were obtained partially from the Stanford
Network Analysis Platform (SNAP) Amazon product metadata dataset and
partially directly from the the Amazon.com website. The product ID's
and categorizations were obtained from the SNAP dataset and the product
descriptions were obtained directly from the website.

We were able to deduce the structure of the hierarchy for the Amazon.com
products directly since all ancestors in the hierarchy were included
with each category label. For example, {}``DVD / Genres / Science
Fiction \& Fantasy / Classic Sci-Fi'' is a single product category
for the DVD, {}``The Time Machine''. Each product was labeled with
multiple categories.

The vocabulary for this experiment was created by including the most
frequent 30K words omitting stopwords.

The two main methods of evaluation for the model are prediction and
topic quality. We compare model performance against 3 similar models
to demonstrate that each component of the model is important for performance.
Specifically, we evaluate models including independent regressors
+ sLDA (hierarchical constraints on labels ignored), HSLDA fit by
running LDA first then running tree-conditional regressions, and HSLDA
fit with fixed random regression parameters.

\subsubsection{Prediction}

Given the expectation o

\subsubsection{Topic Quality and Character}
